Holomorphic field theories and higher algebra

### 2022-09-23 draft.tex

The action functional is
\[
S_{QCD} = \int B\, F^{0,2}_A + \int R \,\dbar_A Q + \int r \,\dbar_A q,
\]
where $\dbar_A = \dbar + A$ is the covariant $\dbar$-operator. 
We will denote this theory by $\cT(N,F)$.\footnote{Although this perturbative theory only depends on the Lie algebra of $G$, the group structure will be relevant when we discuss local operators and moduli of vacua.} 
\end{dfn}

@@

Note that the lower bound on $N$ and $\check{N}$ is simply to ensure that we have nontrivial gauge groups.
%In this case the auxiliary vector space $Z$ is nontrivial (but transforms trivially under the gauge group as in the definition). 
%Our notation is $q,\Tilde{q}$ for elements of the vector spaces $V, \Bar{V}$ and $m\in Z$.
%The magnetic theory is
%\[
%\cT_{\sfM} (N_c, N_f) \define \cT\big(\SU(N_f - N_c) \, , \, (\CC^{N_f - N_c})^{\oplus N_f} \, , \, \CC^{\oplus N^2_f} \, , \, W = m q \Tilde{q} \big)
%\]
%In the definition of $W = m q \Tilde{q}$ we are using the natural contractions which we can describe explicitly as follows.
%Choose a basis $\{q_a^i\}$ for $V = (\CC^{N_f - N_c})^{\oplus N_f}$, $a = 1,\ldots, N_f - N_c$, $i=1,\ldots N_f$ and a dual basis $\{\Tilde{q}_i^a\}$ for $\Bar{V}$.
%Also, let $\{m_i^j\}$ be a basis for $\CC^{\oplus N_f^2}$. 
%Then the cubic polynomial $W$ is
%\[
%W = \sum_{i,j,a} m^i_j q_i^a \Tilde{q}^i_a .
%\]

@@

@@

Many stacks of interest to geometers and representation theorists appear naturally from holomorphic field theories.
For example, the holomorphic $\sigma$-model from $\CC^d$ into a target complex manifold $X$ has
\[
\cM_{\rm vac} \cong T^*[1-d]X.
\]
(If we consider the underlying underived stack, it is just $X$, unless $d = 1$.)
For supersymmetric gauge theories, the moduli stack of the holomorphic twist has the form
\[
\cM_{\rm vac} \cong V/G
\]
where $V$ is a (graded) $G$-representation and $V/G$ denotes the quotient stack.
The factor $V$ depends on the non-gauge fields (i.e., matter multiplets), which are sections of some associated vector bundle;
the quotient by $G$ arises from the gauge transformation.
For example, pure $\cN =1$ holomorphic Yang-Mills theory for group $G$ (i.e., the vector multiplet) has \owen{true?}
\[
\cM_{\rm vac} \cong \fg^*/G,
\]
the coadjoint quotient stack of the group.
(We are working in a $\ZZ/2$-graded world here.)
For general supersymmetric field theories, which involve both a $\sigma$-model and gauge fields, the holomorphic twist has a moduli stack of classical vacua of the form
\[
\cM_{\rm vac} \cong X/G
\]
where $X$ is a derived stack with a $G$-action and $X/G$ denotes the quotient stack.

In studying questions about the classical vacua of supersymmetric field theories, however, physicists have not typically worked with the {\em stack}.
Instead, they have worked with a GIT quotient.
For instance, if the moduli stack of classical vacua were $V/G$, it means physicists work with
\[
V\Sslash G = \Spec(\cO(V)^G),
\]
the scheme whose ring of functions are the $G$-invariant functions on the representation~$V$.
This affine scheme is known as the {\em affinization} of the stack $V/G$, as it is the best affine approximation to the stack (i.e., apply the left adjoint to the inclusion of affine schemes among stacks).
We will call this GIT quotient the moduli {\em space} of classical vacua and denote it by~$M_{\rm vac}(\cT)$.
(Note that the font of M changes between stack and space.)

@@

For the holomorphic QCD theory $\cT(F,N)$ without mesons, 
the derived stack of classical vacua is 
\[
\cM_{\rm vac}(F,N) \cong T^*[1] \left(V_Q \oplus V_Q^*/SL(N)\right).
\]
By contrast the moduli space $M_{\rm vac}$ is a variety already well-studied in classical algebraic geometry.
Before we jump to the full space, let's examine a simpler case involving just the quarks.
A vacuum for the quark field is an element of $V_Q = \Hom(\CC^F, \CC^N)$,
and it is a classic result of invariant theory that
\[
\Hom(\CC^F, \CC^N)\Sslash SL(N) \cong \widehat{{\rm Gr}(F, N)} \subset \Lambda^N \CC^F
\]
where ${\rm Gr}(F, N)$ denotes the Grassmannian of $N$-planes in $\CC^F$ and the hat indicates we take the affine cone with respect to this standard Pl\"ucker embedding.
For the full problem, we have both quarks and antiquarks, so
\begin{align*}
M_{\rm vac}(F,N) 
&\cong V_Q \oplus V_Q^*\Sslash SL(N) \\
&= \Hom(\CC^F, \CC^N) \oplus \Hom(\CC^N, \CC^F)\Sslash SL(N),
\end{align*}
whose ring of functions is also computed by the fundamental theorems of invariant theory.
Those theorems tell us that the GIT quotient is a closed conical subvariety
\[
M_{\rm vac}(F,N)  \hookrightarrow \Lambda^N \CC^F \times \Lambda^N (\CC^F)^* \times \End(\CC^F)
\]
where a pair of matrices $(X,Y) \in \Hom(\CC^F, \CC^N) \times \Hom(\CC^N, \CC^F)$ goes to $(\Lambda^N X, \Lambda^N Y, YX)$, with $\Lambda^N X$ denoting the wedge product of the rows of $X$ (and of columns for $Y$).
The image thus satisfies both sets of Pl\"ucker relations, the condition that the composite $YX$ has rank at most $N$, and some further relations, given by the Second Fundamental Theorem.
\owen{I'm trying to find a completely careful and explicit statement of the $SL(N)$ case.}
In other words,
\[
M_{\rm vac}(F,N) \hookrightarrow \widehat{{\rm Gr}(F, N)} \times \widehat{{\rm Gr}(F, N)} \times {\rm R}(F,N)
\]
with ${\rm R}(F,N) \subset \End(\CC^F)$ the matrices with rank bounded above by~$N$.
The moduli space for the magnetic theory is a bit more complicated but follows from invariant theory as well.

>>>>

The action functional is
\[
S_{QCD} = \int B\, F^{0,2}_A + \int R \,\dbar_A Q + \int r \,\dbar_A q,
\]
where $\dbar_A = \dbar + A$ is the covariant $\dbar$-operator.
%We will denote this theory by $\cT(N,F)$.
\footnote{Although this perturbative theory only depends on the Lie algebra of $G$, the group structure will be relevant when we discuss local operators and moduli of vacua.} 
\end{dfn}

@@

Note that the lower bound on $N$ and $\check{N}$ is simply to ensure that we have nontrivial gauge groups.

%In this case the auxiliary vector space $Z$ is nontrivial (but transforms trivially under the gauge group as in the definition). 
%Our notation is $q,\Tilde{q}$ for elements of the vector spaces $V, \Bar{V}$ and $m\in Z$.
%The magnetic theory is
%\[
%\cT_{\sfM} (N_c, N_f) \define \cT\big(\SU(N_f - N_c) \, , \, (\CC^{N_f - N_c})^{\oplus N_f} \, , \, \CC^{\oplus N^2_f} \, , \, W = m q \Tilde{q} \big)
%\]
%In the definition of $W = m q \Tilde{q}$ we are using the natural contractions which we can describe explicitly as follows.
%Choose a basis $\{q_a^i\}$ for $V = (\CC^{N_f - N_c})^{\oplus N_f}$, $a = 1,\ldots, N_f - N_c$, $i=1,\ldots N_f$ and a dual basis $\{\Tilde{q}_i^a\}$ for $\Bar{V}$.
%Also, let $\{m_i^j\}$ be a basis for $\CC^{\oplus N_f^2}$. 
%Then the cubic polynomial $W$ is
%\[
%W = \sum_{i,j,a} m^i_j q_i^a \Tilde{q}^i_a .
%\]

@@

\owen{Do we want $F > N+1$? The physicists break things into several cases. In particular, I think the moduli is unchanged under quantization only if $F$ is big enough relative to $N$ and $\check{N}$.}

@@

Many stacks of interest to geometers and representation theorists appear naturally from holomorphic field theories.
For example, the holomorphic $\sigma$-model from $\CC^d$ into a target complex manifold $X$ has
\[
\cM_{\rm vac} \cong T^*[1-d]X.
\]
(If we consider the underlying underived stack, it is just $X$, unless $d = 1$.)
For supersymmetric gauge theories, the moduli stack of the holomorphic twist has the form
\[
\cM_{\rm vac} \cong V/G
\]
where $V$ is a (graded) $G$-representation and $V/G$ denotes the quotient stack.
The factor $V$ depends on the non-gauge fields (i.e., matter multiplets), which are sections of some associated vector bundle;
the quotient by $G$ arises from the gauge transformation.
For example, pure $\cN =1$ holomorphic Yang-Mills theory for group $G$ (i.e., the vector multiplet) has
\[
\cM_{\rm vac} \cong \fg^*/G,
\]
the coadjoint quotient stack of the group.
(We are working in a $\ZZ/2$-graded world here.)
Using the invariant pairing on $\fg$ we can further identify this with the adjoint quotient stack~$\fg / G$. 
For general supersymmetric field theories, which involve both a $\sigma$-model and gauge fields, the holomorphic twist has a moduli stack of classical vacua of the form
\[
\cM_{\rm vac} \cong X/G
\]
where $X$ is a derived stack with a $G$-action and $X/G$ denotes the quotient stack.

In studying questions about the classical vacua of supersymmetric field theories, however, physicists have not typically worked with the {\em stack}.
Instead, they have worked with a GIT quotient.
For instance, in the setting of a gauge theory with moduli stack of classical vacua $V/G$, it means they work with 
\brian{I don't like this notation, its confusing since it looks like symplectic reduction which has already appeared. 
What about $[V/G]$?}
\[
V\Sslash G = \Spec(\cO(V)^G),
\]
the scheme whose ring of functions are the $G$-invariant functions on the representation~$V$.
This affine scheme is known as the {\em affinization} of the stack $V/G$, as it is the best affine approximation to the stack (i.e., apply the left adjoint to the inclusion of affine schemes among stacks).
We will call this GIT quotient the moduli {\em space} of classical vacua and denote it by~$M_{\rm vac}(\cT)$.
(Note that the font of M changes between stack and space.)

@@

For the holomorphic QCD theory with $F$ flavors and $N$ colors (without mesons), 
the moduli stack of classical vacua is 
\[
\cM_{\rm vac}(F,N) \cong T^*[1] \left(V_Q \oplus V_Q^*/SL(N)\right).
\]
By contrast the moduli space $M_{\rm vac}$ is a variety already well-studied in classical algebraic geometry.
Before we jump to the full space, let's examine a simpler case involving just the quarks.
A vacuum for the quark field is an element of $V_Q = \Hom(\CC^F, \CC^N)$,
and it is a classic result of invariant theory that
\[
\Hom(\CC^F, \CC^N)\Sslash SL(N) \cong \widehat{{\rm Gr}(F, N)} \subset \Lambda^N \CC^F
\]
where ${\rm Gr}(F, N)$ denotes the Grassmannian of $N$-planes in $\CC^F$ and the hat indicates we take the affine cone with respect to this standard Pl\"ucker embedding.
For the full problem, we have both quarks and antiquarks, so
\begin{align*}
M_{\rm vac}(F,N) 
&\cong V_Q \oplus V_Q^*\Sslash SL(N) \\
&= \Hom(\CC^F, \CC^N) \oplus \Hom(\CC^N, \CC^F)\Sslash SL(N),
\end{align*}
whose ring of functions is also computed by the fundamental theorems of invariant theory.
Those theorems tell us that the GIT quotient is a closed conical subvariety
>>>>>>> 46317a9ff2fa2a1a491852a8c6266ecf16d084c0
\[
M_{\rm vac}(F,N)  \hookrightarrow \Lambda^N \CC^F \times \Lambda^N (\CC^F)^* \times \End(\CC^F)
\]
where a pair of matrices $(X,Y) \in \Hom(\CC^F, \CC^N) \times \Hom(\CC^N, \CC^F)$ goes to $(\Lambda^N X, \Lambda^N Y, YX)$, with $\Lambda^N X$ denoting the wedge product of the rows of $X$ (and of columns for $Y$).
The image thus satisfies both sets of Pl\"ucker relations, the condition that the composite $YX$ has rank at most $N$, and some further relations, given by the Second Fundamental Theorem.
\owen{I'm trying to find a completely careful and explicit statement of the $SL(N)$ case.}
In other words,
\[
M_{\rm vac}(F,N) \hookrightarrow \widehat{{\rm Gr}(F, N)} \times \widehat{{\rm Gr}(F, N)} \times {\rm R}(F,N)
\]
with ${\rm R}(F,N) \subset \End(\CC^F)$ the matrices with rank bounded above by~$N$.
The moduli space for the magnetic theory is a bit more complicated but follows from invariant theory as well.
